---
title: "Thrishul Final Project"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}
library(flexdashboard)
library(ggplot2)
library(dplyr)
library(tidyr)
library(leaflet)


airbnb_data <- read.csv("C:/Users/Sunny/Documents/threads/air-bnb-listings.csv", sep = ";")
head(airbnb_data)
```


```{r}

colnames(airbnb_data) <- make.names(tolower(gsub(" ", "_", colnames(airbnb_data))))

# Handle missing values for 'room_price'
airbnb_data <- airbnb_data %>%
  filter(!is.na(room.price) & !is.na(coordinates))

# Since Room.Price is already an integer, we skip the conversion process

# Convert 'minimum_nights' to integer (if it's not already)
airbnb_data$minimum.nights <- as.integer(airbnb_data$minimum.nights)

# Parse 'coordinates' into two separate columns
airbnb_data <- airbnb_data %>%
  separate(coordinates, into = c("latitude", "longitude"), sep = ",", convert = TRUE)

# Remove any duplicates
airbnb_data <- distinct(airbnb_data)

# Filter outliers in 'room_price' if necessary
# For example, you might want to remove listings above the 99th percentile
quantile_limit <- quantile(airbnb_data$room.price, 0.99, na.rm = TRUE)
airbnb_data <- airbnb_data %>%
  filter(room.price <= quantile_limit)

# Check the cleaned data
head(airbnb_data)

```
Airbnb NY Visualization
===================================================================
## Column {.tabset .tabset-fade }

### Room Prices by Room Type {data-width="350"}

```{r}
ggplot(airbnb_data, aes(x = room.type, y = room.price)) +
  geom_boxplot() +
  labs(title = "Room Prices by Room Type",
       x = "Room Type",
       y = "Price") +
  theme_minimal() +
  theme(plot.margin = margin(t = 8, r = 8, b = 8, l = 8)) # Adjust top, right, bottom, left margins as needed

```


Taking a quick look at this box plot, it tells us about the spread and median prices for different types of rooms. Entire homes or apartments have the widest price range and the highest median price, while shared rooms are the cheapest and have less variation in price. The dots outside the main plot show the outliers, which are unusually high or low prices compared to the rest. It's interesting to note that we see quite a few outliers for entire homes and private rooms.


### Distribution of Room Prices


```{r}
library(ggplot2)
library(plotly)

# Define the breaks for the price bins and create labels for them
breaks <- seq(from = 0, to = max(airbnb_data$room.price, na.rm = TRUE), by = 50)
labels <- paste(head(breaks, -1), tail(breaks, -1), sep = "-")

# Create a factor column with the price ranges using the labels
airbnb_data$price_range <- cut(airbnb_data$room.price, breaks = breaks, labels = labels, include.lowest = TRUE, right = FALSE)

# Calculate the count for each price range bin
price_range_counts <- airbnb_data %>%
  group_by(price_range) %>%
  summarise(count = n())

# Join the counts back onto the original data frame
airbnb_data <- left_join(airbnb_data, price_range_counts, by = "price_range")

# Create the ggplot object
p <- ggplot(airbnb_data, aes(x = room.price, fill = price_range)) +
  geom_histogram(binwidth = 50, color = "black") +
  scale_fill_viridis_d(guide = FALSE) +
  labs(title = "Distribution of Room Prices",
       x = "Room Price ($)",
       y = "Count") +
  theme_minimal()

# Convert the ggplot object to a plotly interactive graph, including the count in the tooltip
ggplotly(p, tooltip = c("fill", "count"))
```



### Count of Room Types
```{r}
# Create a bar chart of Room Types
ggplot(airbnb_data, aes(x = room.type)) +
  geom_bar(fill = "cyan", color = "black") +
  labs(title = "Count of Room Types",
       x = "Room Type",
       y = "Count")

```



### Room Price vs. Number of Reviews

```{r}
# Create a scatter plot of Price vs. Number of Reviews
ggplot(airbnb_data, aes(x = number.of.reviews, y = room.price)) +
  geom_point(alpha = 0.5, color = "red") +
  labs(title = "Room Price vs. Number of Reviews",
       x = "Number of Reviews",
       y = "Room Price ($)")


```



### Spatial Distribution of Airbnb Listings

```{r}
library(ggplot2)

ggplot(airbnb_data, aes(x = longitude, y = latitude)) +
  geom_point(aes(color = room.price), alpha = 0.5, size = 1) +
  scale_color_viridis_c() +
  labs(title = "Spatial Distribution of Airbnb Listings",
       x = "Longitude",
       y = "Latitude",
       color = "Price") +
  theme_minimal()
```



### Map of listings

```{r}

# Assuming airbnb_data is loaded and contains 'longitude' and 'latitude' columns
# Create a color palette for room prices
pal <- colorNumeric(palette = "viridis", domain = airbnb_data$room.price)

# Create the leaflet map
leaflet(airbnb_data) %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addCircleMarkers(
    lng = ~longitude, lat = ~latitude,
    color = ~pal(room.price),
    opacity = 1, fillOpacity = 0.8,
    radius = 5,  # Static radius for each marker
    popup = ~paste("Price: $", room.price, "<br>",
                   "Neighbourhood: ", neighbourhood),
    clusterOptions = markerClusterOptions()  # Cluster options enable here
  ) %>%
  addLegend("bottomright", pal = pal, values = ~room.price,
            title = "Room Price",
            opacity = 1)
```

